Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees

### 2023-03-07 script/manuscript_MEE.Rmd
added citations for academic conclusions

  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), and custom species sequence databases derived via species distribution modelling, and temporal filtering. To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants). To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g. iNaturalist), from a region exceeding the study area. For these candidate species, a modelling approach, such as logistic regression, may be used to identify taxa which warrant further exploration e.g. determine their possibility of presence in metabarcoding samples. We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area. Species distribution models examine the ecological conditions associated with the known occurrence of a species to identify suitable habitats in the study area. This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers. This approach can significantly reduce processing time, increasing the a projects efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.

@@

  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts. 
  Difficulties in temporal models related to variability in drivers of flowering phenology, and relatively small appropriate sample sizes per species.

Based on our results, we believe the overall results between observations of queen Bumble Bee and barcoding are largely congruent. Which relaxes concerns regarding differences between the insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence). This indicates that results from fine resolution survey's, may serve as calibration standards to projects which expand geographically outward from them, and seek to make inference to larger areas.  We analyzed pollen loads from all of the most common bumble bee species in the area and were able to ... (@pyke1982local). The species which appeared most often in loads via the molecular records, yet seldom noted via observations, mostly had affinities to wetlands or stream banks. These habitats are not often sampled by terrestrial ecologists due to logistics and concerns regarding degrading them. This indicates that metabarcoding may be applied in studies where areas, or administrative units, prohibit the disturbances associated sampling.

@@

Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Mertensia* on a per visit basis... General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persist with *Bombus* studies...

  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble). Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*. Historic livestock grazing was often associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle. In particular, the removal of locoweeds (Fabaceae: *Astragalus* L. & *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the United States Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism). Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks for significant periods of time, all processes which lower the water tables and reduced the extent of stream-associated [riverine] wetlands and the mesic meadows fringes which provide habitat for many species of tall *Mertensia* (Boraginaceae, e.g. *M. ciliata* Torr. G. Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L. (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos). Fire suppression further resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading). Finally the effects of Nitrogen deposition, especially given the West's rapidly growing population still pose adverse effects on the abundance of a variety of species of Fabaceae at Urban-Rural interfaces (see @stevens2018atmospheric, @fenn2003ecological). Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing. The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L., using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).

>>>>

  Herein we have resolved major components of the problems of accurately and effectively identifying plant material without diagnostic morphological character states using the A353 Hyb-Seq probes (@johnson2019universal), within a framework which utilizes custom species sequence databases derived via species distribution modelling, and temporal filtering. To increase the accuracy and efficiency of metabarcoding results in plants, we are proposing reducing the number of possible candidate species by generating a user specific databases relevant to the region of study and the ecological characteristics of interest (@bell2022plants). To achieve this goal, we first create a regional list of candidate species using digital collections gleaned from herbaria, survey work, and citizen science (e.g. iNaturalist), from a region exceeding the study area. For these candidate species, a modelling approach, such as logistic regression, may be used to identify taxa which warrant further exploration e.g. determine their possibility of presence in metabarcoding samples. We then use species distribution models to create potential distribution maps for the candidate species to limit the impact of spatial and taxonomic biases in the species list and account for spatial variations in niche availability throughout the study area. Species distribution models examine the ecological conditions associated with the known occurrence of a species to identify suitable habitats in the study area. This approach has the benefit of greatly reducing the size of a sequence database, which allows for the use of genomic data on personal computers. This approach can significantly reduce processing time, increasing the a projects efficiency, particularly as most next-generation sequence data is deposited as raw-sequence reads.

@@

  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts.

Our results indicate the overall information gleaned from observations of queen Bumble Bee foraging and analysis of pollen records are largely congruent. Relaxing concerns regarding differences between the broad insights gleaned from observational, as compared to data derived from the pollen records (@barker2021pollen, @zhao2019topology, @alarcon2010congruence). In general when interaction networks are considered at coarse levels, such as the duration of a season, our perceptions regarding the generality of interactions at smaller time scales may be inflated relative to the actualized interactions within them, e.g. a week (@caradonna2020temporal). These results indicate a possibility that at even finer levels *Bombus* display high amounts of floral fidelity within foraging bouts, an observation which implies that part of the reason for the high efficiency of *Bombus* as a pollinator might partially be related to their lack of movement of hetero-specific pollen (@brosi2013single, @ashman2013toward, @galloni2008visitor, @brosi2016pollinator). The mechanisms behind this observed fidelity are likely related to pollen nutritional values, specifically high concentrations of protein, and the absence of particular amino acids required for larval development in other flower more commonly used by workers (@genissel2002influence, @tasei2008nutritive, @goulson2005causes, @goulson2008diet, @hanley2008breeding).

@@

  Some foraging preferences of *Bombus*, both at this field site and across a great many localities globally emerge from this work, which reiterates the needs for land managers to maintain relatively high amounts of members of the famillies Fabaceae, Boraginaceae, and Ranunculaceae, in Western North American montane landscapes (@goulson2005causes, @goulson2010bumblebees, @liang2021evolutionary, @bontvsutvsnaja2021bumble). Numerous historic, and some ongoing, land management practices reduce the ability of many landscapes to support stable populations of *Bombus*. Historic livestock grazing was often associated with the targeted removal of many species of plants which are known to have compounds toxic to cattle. In particular, the removal of locoweeds (Fabaceae: *Astragalus* L. & *Oxytropis* DC.) and larkspurs (Ranunculaceae: *Delphinium*) were common across public lands administered by the United States Forest Service (@ralphs1988herbicide, @aldous1919eradicating, @ralphs2003mechanism). Further actions, generally initiated by early settlers, involved the channelization and incising of streams, culling of beavers, and leaving cattle concentrated on higher order stream banks for significant periods of time, all processes which lower the water tables and reduced the extent of stream-associated [riverine] wetlands and the mesic meadows fringes which provide habitat for many species of tall *Mertensia* (Boraginaceae, e.g. *M. ciliata* Torr. G. Don.) widely distributed across Western North America, and to an extent *Delphinium barbeyi*  and many species of native *Trifolium* L. (@dahl1990wetlands, @naiman1988alteration, @belsky1999survey, @cooke1976arroyos). Fire suppression further resulted in the succession of many Aspen (*Populus tremuloides* Michx.) groves to Conifer stands, decreasing the mosaic of age structured habitats in many landscapes, adversely effects habitat for tall *Mertensia* species and several species of *Delphinium* (@brewen202176, @keane2002cascading). Finally the effects of Nitrogen deposition, especially given the West's rapidly growing population still pose adverse effects on the abundance of a variety of species of Fabaceae at Urban-Rural interfaces (see @stevens2018atmospheric, @fenn2003ecological). Current solutions to these issues, involve targeted burns, reintroduction of beavers and beaver habitat analogs, and the possibility of re-seeding a variety of 'locoweeds' and 'larkspurs' in areas now seldom used, or only used for early, grazing. The highly enthusiastic response of land managers, and homeowners, to plant *Ascelpias* L., using genetically appropriate materials, to improve Monarch Butterfly (*Danaus plexippus* L.) habitat provides an effective framework for the latter (@oberhauser2015monarchs, @basey2015producing).
